---
title: "Exploratory data analysis - An example on TITANIC dataset"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data importing

Import the data into `r` by using function `read.csv`, we use the parameter `na.string` to specify the pattern of missing values.
```{r}
df = read.csv('titanic_train.csv', na.strings = c("NA", "NaN", "", " "))
```

You also see some properties of dataset with `dim()` and `str()`.
```{r}
dim(df)
str(df)
```

The `dim()` function gives us a vector of 2 items including the number of rows/records/observation and the number of columns/features/fields. Meanwhile, function `str()` shows the basic characteristics of each columns such as data type and some values in the column. 

Some common statistics are also calculated by using `summary()` function:
```{r}
summary(df)
```

The function `object.size()` returns the memory allocation of data frame. 
```{r}
object.size(df)
```

To reduct the size of data frame, we can specify the data type of each column after reading:
```{r}
print(paste('Size of df before optimize:', object.size(df)))
df$survived = as.logical(df$survived)
df$ticket = as.numeric(df$ticket)
print(paste('Size of df after optimize:', object.size(df)))
```


## Data cleaning
### Detecting missing values
In `R`, function `is.na()` return the logical vector which indicate if the value is missing or not.
Count the number of missing values in the dataset:
```{r}
sum(is.na(df))
```
To count missing values per columns:
```{r}
colSums(is.na(df))
```

### Handling missing values
Drop the columns:
```{r}
df <- subset(df, select=-c(age, cabin))
print(dim(df))
```

Drop the rows:
```{r}
df <- subset(df, select=!is.na(df)$embarked)
```


### Detecting the outliers using 

